---
title: "Homework 1 - Statistical Modelling and Inference"
author: "Ván Bálint"
date: "October 7, 2015"
output: html_document
---

## 1.

IFF $x^{T}X^{T}Xx>=0$ for all x, than $X^{T}X$ is positive semi definite.

$x^{T}X^{T}Xx=(Xx)^{T}Xx$

$Xx$ is a $p*1$ vector, so $(Xx)^{T}Xx$ is an inner product.

An inner product can never be negative, so $X^{T}X$ is always positive semi definite.

## 2.
```{r, echo=FALSE}
download.file("http://statweb.stanford.edu/~tibs/ElemStatLearn/datasets/prostate.data", "prostate.txt")
prostate<-read.table("prostate.txt")
prostate[,1:8] <- scale(prostate[,1:8],TRUE,TRUE)
```

####Correlations of predictors in the prostate cancer data.
```{r, echo=FALSE}
cor(prostate[,1:8])
```

####Linear model fit to the prostate cancer (train) data.
```{r, echo=FALSE}
lm_prost1 <- lm(lpsa ~ lcavol + lweight + age + lbph + svi + lcp + gleason + pgg45, data=prostate[prostate$train==TRUE,])
summary(lm_prost1)
lm_prost2 <- lm(lpsa ~ lcavol + lweight + lbph + svi, data=prostate[prostate$train==TRUE,])
RSS1 <- sum(residuals(lm_prost1)^2)
RSS2 <- sum(residuals(lm_prost2)^2)
Fvalue <- (RSS2-RSS1)/(9-5)/RSS1*(67-9)
```
The F-value is `r Fvalue`, with the corresponding p-value of `r pf(Fvalue,4,58,lower.tail=FALSE)`.

The mean prediction error on the test data is `r mean(abs(predict.lm(lm_prost2, prostate[prostate$train==FALSE,]) - prostate$lpsa[prostate$train==FALSE]))`. In contrast, prediction using the mean training value of lpsa ha a test error of `r mean(abs(mean(prostate$lpsa[prostate$train==TRUE]) - prostate$lpsa[prostate$train==FALSE]))`.

Lcp and lcavol have a strong correlation. When this is the case their coefficients can behave in unexpected ways if both are correlated to the dependent variable.

## 3.

1.

$H=\phi(\phi^{T}\phi)^{-1}\phi^{T}$

$H^{T}=(\phi(\phi^{T}\phi)^{-1}\phi^{T})^{T}=(\phi^{T})^{T}((\phi^{T}\phi)^{-1})^{T}\phi^{T}=\phi(\phi^{T}\phi)^{-1}\phi^{T}=H$

2.


